---
title: "Lab 05- Wrangling spatial data"
author: "Cat Seitz"
date: "01/31/2023"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1

```{r filter_dennys_for_alaska}

dn<-dennys

dn_ak <- dn %>%
  filter(state == "AK")
nrow(dn_ak)

```
There are 3 Denny's locations in Alaska. 

### Exercise 2

```{r filter_lq_for_alaska}

lq<-laquinta

lq_ak <- lq %>%
  filter(state == "AK")
nrow(lq_ak)

```

There are 2 La Quinta locations in Alaska. 

### Exercise 3

We need to calculate 6 distances between the Alaska establishment locations. 

```{r join_dataframes}

dn_lq_ak <- full_join(dn_ak, lq_ak, by = "state")
dn_lq_ak

```

### Exercise 4

There are 6 observations in our joined data frame, which aligns with what we want. The variables are the addresses of the Denny's and La Quinta locations, as well as their longitudes and latitudes. The data frame is organized so each row is a pair of Denny's and La Quinta locations, with none repeating. 

### Exercise 5

```{r haversine}

haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}

```



### Exercise 6

I added a column to the Denny's and La Quinta's in Alaska data frame for the distance between each pair of locations. 

```{r calc_distance}

dn_lq_ak$distance <-
  haversine(dn_lq_ak$longitude.x,
            dn_lq_ak$latitude.x,
            dn_lq_ak$longitude.y,
            dn_lq_ak$latitude.y,
            3)

```


### Exercise 7

```{r minimum_distance}

dn_lq_ak_mindist <- dn_lq_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

```

### Exercise 8

For each Denny's location in Alaska, there is a La Quinta within 6 kilometers. 

```{r visualization}
dn_lq_ak_mindist %>%
  ggplot (aes(x=closest))+
  geom_histogram(binwidth = 3)

```


### Exercise 9 - North Carolina

```{r filter_NC}

#filter dennys
dn_nc <- dn %>%
  filter(state == "NC")

#filter la quinta
lq_nc <- lq %>%
  filter(state == "NC")

#join data frames
dn_lq_nc <- full_join(dn_nc, lq_nc, by = "state")

#calculate distance
dn_lq_nc$distance <-
  haversine(dn_lq_nc$longitude.x,
            dn_lq_nc$latitude.x,
            dn_lq_nc$longitude.y,
            dn_lq_nc$latitude.y,
            3)

#find minimum distance
dn_lq_nc_mindist <- dn_lq_nc %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

#plot histogram
dn_lq_nc_mindist %>%
  ggplot (aes(x=closest))+
  geom_histogram(binwidth = 10)

```





### Exercise 10 - Texas

```{r filter_tx}

#filter dennys
dn_tx <- dn %>%
  filter(state == "TX")

#filter la quinta
lq_tx <- lq %>%
  filter(state == "TX")

#join data frames
dn_lq_tx <- full_join(dn_tx, lq_tx, by = "state")

#calculate distance
dn_lq_tx$distance <-
  haversine(dn_lq_tx$longitude.x,
            dn_lq_tx$latitude.x,
            dn_lq_tx$longitude.y,
            dn_lq_tx$latitude.y,
            3)

#find minimum distance
dn_lq_tx_mindist <- dn_lq_tx %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

#plot histogram
dn_lq_tx_mindist %>%
  ggplot (aes(x=closest))+
  geom_histogram(binwidth = 5)

```

### Exercise 11 - California

```{r filter_CA}

#filter dennys
dn_ca <- dn %>%
  filter(state == "CA")

#filter la quinta
lq_ca <- lq %>%
  filter(state == "CA")

#join data frames
dn_lq_ca <- full_join(dn_ca, lq_ca, by = "state")

#calculate distance
dn_lq_ca$distance <-
  haversine(dn_lq_ca$longitude.x,
            dn_lq_ca$latitude.x,
            dn_lq_ca$longitude.y,
            dn_lq_ca$latitude.y,
            3)

#find minimum distance
dn_lq_ca_mindist <- dn_lq_ca %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))

#plot histogram
dn_lq_ca_mindist %>%
  ggplot (aes(x=closest))+
  geom_histogram(binwidth = 10)

```

### Exercise 12

















